---
title: "PSYC 30550"
author: "Sooyoun Zong"
short_title: "Ses-infer project pilot RMD"
---

## HW (15) Analysis Plan

```{r plan, echo = TRUE}

1. Independent variable Vignettes (two explanation levels X two social classes)
- 3 higher-class topic vignettes + 3 lower-class topic vignettes
- Each vignette has basic (detailed explanation) and neutral version (no explanation)

2. Dependent variable
- Listener’s knowledge rating (7-point Likert)
- Listener's social class rating (10-point Likert)
- Participant’s social class + knowledge of various topics

3. Analysis

A. Relationship between participant social class and knowledge
- Analysis: Correlation analysis for manipulation check ∙ for selecting topics for future study
- Plot: Scatter plot + fitted line visualizing the relationship
- Table: Correlation matrix with correlation coefficients, significance levels

B. Compare knowledge and social class ratings between conditions
- Analysis: ANOVA (among four groups) or Independent samples t-test (between two explanation conditions) to compare (1) average knowledge ratings (2) social class ratings
- Plots: (1) Histograms visualizing the data distribution (2) Bar graph showing the distribution of ratings for each condition
- Tables: (1) Descriptive statistics of means & standard deviations (2) Summary of ANOVA / t-test results

```

# HW (16) -1 Example correlation analyses

```{r descriptive statistics, echo = TRUE}

**Data not collected yet**

library(dplyr)
library(ggplot2)

data %>%
  summarise(
    avg_participant_social_class = mean(participant_social_class, na.rm = TRUE),
    sd_participant_social_class = sd(participant_social_class, na.rm = TRUE),
    sd_knowledge = sd(knowledge, na.rm = TRUE),
    avg_social_class = mean(social_class, na.rm = TRUE),
    sd_social_class = sd(social_class, na.rm = TRUE))

```

```{r hypothesis testing, echo = TRUE}

**Data not collected yet**
  
library(psych)

cor.test(data$participant_social_class, data$topic_knowledge)

ggplot(data, aes(x = participant_social_class, y = topic_knowledge)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Relationship between participant social class and knowledge of topic",
       x = "Participant social class",
       y = "Knowledge of topic")

cor_matrix <- cor(data[c("participant_social_class", "topic_knowledge")], use = "complete.obs")
print(cor_matrix)

```

# HW (16) -2 Example advanced analysis (mixed-effects model)

```{r hypothesis testing, echo = TRUE}

**Data not collected yet**
  
library(lme4)

model <- lmer(knowledge ~ vignette_type + (1 | participant_id), data = data)
summary(model)

# knowledge is predicted by vignette_type. Random intercepts for each participant_id accounts for the repeated measures (necessary?).

```

### HW (17) In-line code references
**Data not collected yet**

The study explores the relationship between participants' social class and their knowledge of various conversation topics. The social class of r nrow(data) individuals was measured on a 10-point likert scale (MacArthur Subjective SES ladder), and their knowledge across different topics (that are typically well-known by either higher or lower social class group) was assessed on a 7-point likert scale.

The analyses will reveal an average participant social class of r mean(data$participant_social_class, na.rm = TRUE), with a standard deviation of r sd(data$participant_social_class, na.rm = TRUE). Knowledge on the topics will show an average score of r mean(data$topic_knowledge, na.rm = TRUE), and a standard deviation of r sd(data$topic_knowledge, na.rm = TRUE).

The correlation test will indicate a significant relationship between social class and knowledge, with a correlation coefficient of r cor_test_result$estimate and a p-value of r cor_test_result$p.value. This will suggest that as participants' social class increases, their knowledge on higher-class common ground topics will also tend to increase, and vice versa.

A scatter plot will visualize this relationship with a fitted line indicating the direction and strength of the relationships. The correlation matrix will additionally reveal a coefficient of r cor_matrix["participant_social_class", "topic_knowledge"].

The findings will ultimately suggest that socioeconomic factors may indeed play a role in shaping individuals' cultivated knowledge of specific culture amd experiences.

