---
title: "PSYC 30550"
author: "Sooyoun"
short_title: "Ses-infer project Rmd"
---

# HW 4,5
```{r load-function, echo=TRUE}
source("script.R")  # Replace with the actual path to your R script
hello_world("My friend", FALSE, 10, 3)
```

# Nullam sed sapien sit amet nulla tincidunt aliquam.
## Phasellus in turpis gravida, cursus felis vel, eleifend enim.
### Donec non metus porttitor, gravida dui a, pretium eros.

- Nullam sed sapien sit amet nulla tincidunt aliquam.
- Phasellus in turpis gravida, cursus felis vel, eleifend enim.
- Donec non metus porttitor, gravida dui a, pretium eros.

1. Nullam sed sapien sit amet nulla tincidunt aliquam.
2. Phasellus in turpis gravida, cursus felis vel, eleifend enim.
3. Donec non metus porttitor, gravida dui a, pretium eros.

[Lorem Ipsum](https://www.lipsum.com/)

<!-- The weather outside is crazy -->

# HW 6

```{r data-read-in}
library(readxl)
data <- read_xlsx("dataset.xlsx")

# Examining the dataset structure (Tidiness check)
str(data)

# Columns are single variables.
# Rows are single observations.
# Cells are single measurements.
# My dataset appears to be tidy.

```

```{r create-intermediate-dataset}
library(dplyr)
library(writexl)

# Stimuli type, reports of perceived ses, perceived warmth and competence, perceiver ses are necessary variables for my analyses.
data_prep <- data %>%
  select(type, age, edu, class, rich, hire, a_inc, a_edu, a_class, comp, warm) %>%
  filter(age > 35)

data_prep$type <- as.factor(data_prep$type)

# Writing the intermediate dataset
write_xlsx(data_prep, "data_prep.xlsx")

# Tidiness check
str(data_prep)

```


